\chapter{Preface}
In 1917, the world would be forever changed.
In the Russian Empire, there existed a political party known as the Russian Social Democratic Party, or the Bolsheviks.
Fed up with their empire's economic instability, industrial underdevelopment, appalling treatment of the working-class people, and the lack of initiative by the new provisional government to address these issues, they realized that a ``reformist'' plan of action was not tenable.
In October that year, they overthrew the liberal provisional government and established a new nation, Soviet Russia.
Over the next five years, they fought against the fascist forces of the White Army and its international allies to secure their massive country, and in 1922, such security was permanently recognized.
The Union of Soviet Socialist Republics, otherwise known as the Soviet Union, became the successor to the autocratic, capitalist badlands of the Russian Empire.
The Soviet Union would lead its working people on the path towards communism, a state of society in which there is no existence of class, money or the state, and the means of production are commonly owned by all working members of the community. 
 
The establishment of the Soviet Union would inspire many similar revolutions across the world, in which vanguard parties of proletariats (the working class) would take power in their capitalist-torn nations and established socialist societies.
Power over the means of production would be taken from the greedy bourgeoisies (the ruling class) and given over to the proletariat, so that they could build their nations into industrial powerhouses of modernity, together and democratically. 
 
This way of thinking is commonly known as Marxism-Leninism, named after its founders Karl Marx and Vladimir Lenin, the ladder of whom was the founder of the Soviet Union.
Marxist-Leninist states have been established all across the globe, and the achievements of these nations---both economically and socially---have been tremendous, and have rivalled the accomplishments of their capitalist counterparts.
Unfortunately, they have also been under constant scrutiny and persecution from the imperialist action of said counterparts, and most Marxist-Leninist nations have either submitted to ``free market'' regression and eventual dissolution, or have been forcefully overthrown by strong capitalist powers, most notably the United States of America. 
 
Nevertheless, Marxist-Leninist theory persists among a small number of proletariats today, and this persistence has been accompanied by a general growing disillusionment with capitalism and its unstable forces across the world.
A 2019 study by the PR firm Edelmen shows that 56\% of the world's population is ``dissatisfied'' with capitalism, and an Axios survey done the same year showed that at least 40\% of Americans prefer their varying definitions of ``socialism'' to capitalism. 
 
Clearly, the proletarians of the world are once again beginning to wake up to the true nature of the situations they live in.
And the internet has only made it infinitely easier for the average working-class person to communicate with and learn from other like-minded individuals from any location on earth.
Enter r/communism, an online community powered by the social media website Reddit, and the largest forum on the internet for communists who subscribe to Marxist-Leninist thought.
Established in 2008, r/communism has grown to hold approximately 131,000 members as of March 2020.
In those twelve years---almost exclusively the past five---dozens of large written posts, known as ``masterposts'' or ``megathreads,'' have been created by members of our subreddit on various issues and topics pertaining to our ideology.
Whether they be detailed analyses of former ML states that cite academic sources, or well-written essays refuting the many lies and myths of socialism, there is a lot to be learned here if you are a newly professed ML looking for short answers to the many questions you may have. 
 
This book is a compilation of 54 of said masterposts and megathreads, written by two dozen authors over the course of a near six-year period, from mid-2014 up until just this year.
Some of them have been somewhat altered from their original state to be more suited for book form, but otherwise remain virtually identical to their original posts.
We obviously do not advocate these texts over the works of professional Marxist-Leninist thinkers, but for those who are quite new to the ideology, these relatively simple and straightforward posts can offer a suitable starting point. 
 
Even if you are not a communist or Marxist-Leninist, but are simply intrigued by some of our arguments, we welcome you to read this book as well. Regardless of your ideology, we hope that you find something educational in these pages, and that you can develop a better understanding of who we are, and what we stand for! 
\\\\
\begin{large}
\emph{Long live the struggle against capitalism, imperialism and fascism!}\\\\
\emph{Long live the people!}\\\\
\emph{Workers of the world, unite!}
\end{large}
